---
title: "EDA"
format: html
editor: visual
---

##Introduction

This project explores data from the Diabetes Health Indicators Dataset, based on the BRFSS 2015 survey. The dataset contains self-reported health and lifestyle information from over 250,000 adults in the US. The goal is to create a model for identifying individuals with diabetes using various predictors.

In this exploratory data analysis, we're going to examine 3  variables that could be associated with diabetes status: BMI (Body Mass Index), Physical Activity, and High Blood Pressure (HighBP). I chose these variables because based on a lot of prior research and general knowledge, these variables tend to be strong predictors of diabetes risk. This EDA will help inform which variables may be useful for future modeling. The ultimate goal is to predict the existence of diabetes in an individual based on the variables present in the dataset. 


## Setting all my libraries

```{r}
library(readr)
library(tidyverse)
```



## Data: I'm only converting the three variables into factors that I'm going to be working with because adding other variables requires more computational power that makes the model run for longer. I'm also looking at a glimpse of the dataset and checking if there are any NA values

```{r}
diabetes <- read_csv("diabetes_012_health_indicators_BRFSS2015.csv")

glimpse(diabetes_data)

colSums(is.na(diabetes_data))
```

```{r}
diabetes_data <- diabetes |>
  mutate(
    Diabetes_binary = factor(Diabetes_012, labels = c("No Diabetes", "Prediabetes", "Diabetes")),
    PhysActivity = factor(PhysActivity, labels = c("No", "Yes")),
    HighBP = factor(HighBP, labels = c("No", "Yes"))
  )
```

##Summarizations

### Univariate Summary for BMI: We see that BMI is right-skewed with the median being 27. The mean is slightly higher than the median which explains the right skewed distribution. For the model, I'm assuming or anticipating that a higher BMI will results in a greater risk of diabetes.

```{r}
summary(diabetes_data$BMI)

ggplot(diabetes_data, aes(x = BMI)) + geom_histogram(binwidth = 2, fill = "blue", color = "gray") + labs(title = "BMI Distribution", x = "BMI", y = "Count")
```

### Univariate Summary: Physical Activity. As we can see, majority of the people do participate in physical activity. Howerver, there is still a good number of people that don't partake in physical activity. I'm anticipating that the people who don't partake will show a higher risk of diabetes in the modeling stage. 

```{r}
ggplot(diabetes_data, aes(x = PhysActivity)) + geom_bar(fill = "green") + labs(title = "Physical Activity", x = "Active or Not", y = "Count")
```


### Univariate Summary: High Blood Pressure. We can see that there is a substanial number of peoople reporting a high blood pressure even though the No column has more. A higher blood pressure results in a higher risk of diabetes as well. 

```{r}
ggplot(diabetes_data, aes(x = HighBP)) + geom_bar(fill = "red") + labs(title = "High Blood Pressure", x = "High BP or Not?", y = "Count")
```


### Bivariate Relationship with Diabetes. These three plots now explore the bivariate relationship with diabetes.

In the first plot, we can see that there is a slightly higher BMI associated with people who do have diabetes. People with no diabetes tend to have the lowest BMI. 

```{r}
ggplot(diabetes_data, aes(x = Diabetes_binary, y = BMI, fill = Diabetes_binary)) + geom_boxplot() + labs(title = "BMI by Diabetes", x = "Diabetes", y = "BMI")
```

In this second plot, we can clearly see that that individuals that don't partake in physical activity tend to have more cases of diabetes. Even though there are still people who do physical activity and have diabetes, the proportion of the latter is much higher. 

```{r}
ggplot(diabetes_data, aes(x = PhysActivity, fill = Diabetes_binary)) +geom_bar(position = "fill") +labs(title = "Diabetes Physical Activity", y = "Proportion") 
```

In the last plot, we can see the highest correlation thus far of the three plots. A higher BP tends to lead to higher number of diabetics. This isn't a surprise based on past research and general knowledge. 

```{r}
ggplot(diabetes_data, aes(x = HighBP, fill = Diabetes_binary)) + geom_bar(position = "fill") + labs(title = "Diabetes by High BP", y = "Proportion")
```








